'
Data Mining Principles | Assignment 2
Azizha Zeinita | 12334537
'

```{r}
library("knitr")
library(ggplot2)
library(cluster)
library(factoextra)
library(gridExtra)
library(MASS)

data <- Boston
```


'1. Select the numeric variables that you think are appropriate and useful. 
Use kmeans and Gaussian Mixture models.'

#Select the numeric variables
```{r}
rm=data[,7]
dis=data[,9]
crim=data[,1]
medv=data[,14]
df <- data.frame(rm,dis,crim,medv)

X<- sample(c(rep(0, 0.7 * nrow(df)), rep(1, 0.3 * nrow(df))))
table(X)
train <- df[X == 0, ]  
test <- df[X== 1, ]   
```

'2. Split into train and test (70-30). Scale the data'
```{r}
X.train.mean = colMeans(train)
X.train.sd   = sapply(train, sd)
X.train.scale  = scale(train, center=X.train.mean, scale=X.train.sd)
X.test.scale   = scale(test, center=X.train.mean, scale=X.train.sd) #scaling test by train parameters
```

'3. Generate the K-means solution'
```{r}
vaf = c()
centroid = c()
size = c()
num_clusters = seq(2, 10, by=1)
for (i in num_clusters){
  set.seed(123)
  kmeans_train = kmeans(X.train.scale,centers=i,nstart=70)
  vaf = append(vaf,1 - kmeans_train$tot.withinss / kmeans_train$totss)
  centroid = append(centroid, kmeans_train$centers)
  size = append(size,kmeans_train$size)
}
```

'4. Perform a Scree test | 5. Show the scree plot.'
```{r}
qplot(c(2:10), vaf) +
  geom_line() +
  xlab("# Cluster") +
  ylab("vaf") +
  ggtitle("Elbow Scree Plot") +
  ylim(0.3,1)
```

'6.Choose 1 K-means solution (the best K from the last step) to retain from the many 
solutions that you have generated'
```{r}
kmeans_train_single = kmeans(X.train.scale,centers=5,nstart=70)
kmeans_test_single = kmeans(X.train.scale,centers=kmeans_train_single$centers,nstart=70)

vaf_train = 1 - kmeans_train_single$tot.withinss / kmeans_train_single$totss
centroid_train = kmeans_train_single$centers
size_train = kmeans_train_single$size

vaf_test = 1 - kmeans_test_single$tot.withinss / kmeans_test_single$totss
centroid_test = kmeans_test_single$centers
size_test = kmeans_test_single$size


print(data.frame(vaf_train,vaf_test))
print(data.frame(centroid_train,centroid_test))
print(data.frame(size_train,size_test))
```

'7. Generate 3-5 Gaussian Mixtures (GM) | 8. Choose one solution & do interpretation'
```{r}
require(mclust)
set.seed(123)
gm = Mclust(X.train.scale,G=3:5)
gm$bic #bic value of the selected model
gm$BIC #based on the table shown, the selected model is GMM with number of component=5 and model=VEV
summary(gm) #most of the data are clustered in cluster 3 (51.23%), and the least is in cluster 3 (2.25%)
```

'9. Build a GM model with the best components on train data 
and compare it with the train KMeans solution from an interpretability perspective'
```{r}
p1 = fviz_cluster(kmeans_train_single,geom = "point", data=X.train.scale) + ggtitle("K-Means")
p2 = fviz_cluster(gm,geom = "point", data=X.train.scale) + ggtitle("GM G=5")
grid.arrange(p1, p2, nrow = 2)

kmeans_train_single$centers
gm$parameters$mean
```

'10. Summarize results and interpret the clusters/segments you choose as your final solution.'
```{r}
kmeans_train_single$centers
kmeans_train_single$size
gm$parameters$mean
summary(gm)
```
'All interpretations are available in pdf document'
